---
title: "Chapter 21 - Exercises - R for Data Science"
author: "Francisco Yira Albornoz"
date: "19 de noviembre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 21.2 For loops

## 21.2.1 Exercises

1. Write for loops to:
  
  1. Compute the median of every column in `mtcars`
```{r}
median_mtcars <- vector("double", length(mtcars))
for (i in seq_along(mtcars)) {
  median_mtcars[[i]] <- median(mtcars[[i]])
}
median_mtcars
```
  
  2. Determine the type of each column in `nycflights13::flights`
```{r}
typecol_flights <- vector("character", length(nycflights13::flights))
for (i in seq_along(nycflights13::flights)) {
  typecol_flights[[i]] <- typeof(nycflights13::flights[[i]])
}
typecol_flights
```
  
  3. Compute the number of unique values in each column of `iris`
```{r}
n_unique_val_iris <- vector("integer", length(iris))
for (i in seq_along(iris)) {
  unique_val_iris <- unique(iris[[i]])
  n_unique_val_iris[[i]] <- length(unique_val_iris)
}
n_unique_val_iris
```
  
  4. Generate 10 random normals for each of $\mu = -10$, $0$, $10$, and $100$
```{r}
means <- c(-10, 0, 10, 100)
random_normals <- vector("list", length(means))
for (i in seq_along(means)) {
  random_normals[[i]] <- rnorm(10, mean = means[[i]])
}
random_normals
```
  
2. Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors
  a. 
```{r}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}
out
```

Alternative without for loop:
```{r}
stringr::str_c(letters, collapse = "")
```

  b. 
```{r}
x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))
```

Alternative without for loop:
```{r}
sd(x)
```

  c.
```{r eval = FALSE}
x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
out
```
  
```{r eval = FALSE}
cumsum(x)
```
  
3. Combine your function writing and for loop skills:

  1. Write a forloop that `prints()` the lyrics to the childrenÅfs song ÅgAlice the camelÅh.
```{r}
names_song <- c("Alice", "Ruby", "Sally", "Felix", "Lilly", "Andy", "Larry", "Sammy")
animals_song <- c("Camel", "Rabbit", "Sloth", "Fox", "Ladybug", "Ant", "Lizard", "Spider")
numbers_song <- c("one", "two", "two", "four", "five", "six", "seven", "eight")
what_they_have <- c("hump", "ears", "toes", "legs", "spots", "legs", "stripes", "legs")

for (i in seq_along(names_song)) {
  for (j in 1:4) {
    writeLines(
      stringr::str_c(
        names_song[[i]],
        " the ",
        animals_song[[i]],
        " has ",
        numbers_song[[i]],
        " ",
        what_they_have[[i]],
        "."
      )
    )
  }
  writeLines(stringr::str_c("Go ", names_song[[i]], " go!\n"))
  
}
```
  
  2. Convert the nursery rhyme Ågten in the bedÅh to a function. Generalise it to any number of people in any sleeping structure.
  
```{r}
library(stringr)
library(english)

people_fell_song <-
  function(n_people = 10,
           sleeping_structure = "bed") {
    # Input validation
    if (n_people < 1)
      stop("n_people should be greater than zero")
    
    n_people <- as.integer(n_people)
    if (is.na(n_people))
      stop("n_people should be a valid integer")
    
    
    for (i in 1:n_people) {
      # Turning number of people into a word
      people_in <- as.english(n_people + 1 - i)
      
      # Handling special case for first strophe
      if (i == 1) {
        writeLines("Here we go!")
      } else {
        writeLines(str_c(str_to_title(people_in), "!"))
      }
      
      # Lines that appear in every strophe
      writeLines(str_c("There were ", people_in, " in the ", sleeping_structure))
      writeLines("and the little one said,")
      
      # Handling special case for last strophe
      if (people_in > 1) {
        writeLines('"Roll over, roll over."')
        writeLines("So they all rolled over and one fell out.\n")
      } else {
        writeLines('"Good night!"')
      }
      
    }
  }

people_fell_song(5, "sofa")
```
  
  3. Convert the song Åg99 bottles of beer on the wallÅh to a function. Generalise to any number of any vessel containing any liquid on any surface.
  
```{r}
library(stringr)

song_bottles_beer <- function(n_vessels = 99, vessel = "bottles", liquid = "beer", surface = "wall") {
    # Input validation
    if (n_vessels < 1)
      stop("n_vessels should be greater than zero")
    
    n_vessels <- as.integer(n_vessels)
    if (is.na(n_vessels))
      stop("n_vessels should be a valid integer")
    
    
    for (i in 1:(n_vessels + 1)) {
      # Turning number of bottles into a word
      vessels_left <- (n_vessels + 1 - i)
      
      if (vessels_left >= 1) {
        writeLines(str_c(vessels_left, " ", vessel, " of ", liquid, " on the ", surface, ", ", vessels_left, " ", vessel, " of ", liquid, "."))
      }
      
      if (vessels_left >= 2) {
        writeLines(str_c("Take one down and pass it around, ", (vessels_left - 1), " ", vessel, " of ", liquid, " on the ", surface, ".\n"))
      }
      
      if (vessels_left == 1) {
        writeLines(str_c("Take one down and pass it around, no more ", vessel," of ", liquid," on the ", surface,".\n"))
      }
      
      if (vessels_left == 0) {
        writeLines(str_c("No more ", vessel," of ", liquid," on the ", surface,", no more ", vessel, " of ", liquid, "."))
        writeLines(str_c("Go to the store and buy some more, ", n_vessels," ", vessel," of ", liquid," on the ", surface,"."))
      } 
        
   
      
    }
  }

song_bottles_beer(5, vessel = "glasses", liquid = "water", surface = "table")
```
  
  
4. ItÅfs common to see for loops that donÅft preallocate the output and instead increase the length of a vector at each step:

```
output <- vector("integer", 0)
for (i in seq_along(x)) {
  output <- c(output, lengths(x[[i]]))
}
output

```
How does this affect performance? Design and execute an experiment.

```{r}
library(microbenchmark)

func_wo_preallocate <- function(x) {
  output <- vector("integer", 0)
  for (i in seq_along(x)) {
    output <- c(output, lengths(x[[i]]))
  }
  output
}

func_with_preallocate <- function(x) {
  output <- vector("integer", length(x))
  for (i in seq_along(x)) {
    output[[i]] <- lengths(x[[i]])
  }
 output  
}

test_vec <- 1:1000

microbenchmark(func_wo_preallocate(test_vec),
               func_with_preallocate(test_vec),
               times = 1000L)


```

## 21.3 For loop variations

###  21.3.5 Exercises

1. Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, `files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)`, and now want to read each one with `read_csv()`. Write the for loop that will load them into a single data frame.

```
output <- vector("list", length(files))
for (i in seq_along(files)) {
  output[[i]] <- read_csv(files[[i]])
}

my_big_df <- dplyr::bind_rows(output)
```
2. What happens if you use `for (nm in names(x))` and `x` has no names? What if only some of the elements are named? What if the names are not unique?

```{r}
my_vec <- 1:5
names(my_vec) <- c("a", "a", "b", "c")

for (nm in names(my_vec)) {
  print(my_vec[[nm]])
}
```

If the names are not unique then the subscripting will return the *first element with that name* each time that the duplicated name appears. If only some of the elements are named, there will be an error.

3. Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, `show_mean(iris)` would print:

```
show_mean(iris)
#> Sepal.Length: 5.84
#> Sepal.Width:  3.06
#> Petal.Length: 3.76
#> Petal.Width:  1.20
```
(Extra challenge: what function did I use to make sure that the numbers lined up nicely, even though the variable names had different lengths?)

```{r}

show_mean <- function(x) {
  
  for (i in seq_along(mtcars)) {
    mean_col <- round(mean(mtcars[[i]]), 2)
    
    name_with_colon <- str_c(names(mtcars)[[i]], ":")
    
    name_pad <- str_pad(name_with_colon, width = 6, side = "right")
    
    print(str_c(name_pad, mean_col, sep = " "))
  }
  
}

show_mean(mtcars)

```

4. What does this code do? How does it work?

```
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```
`trans` is a list where each element is a function, designed to do a certain transformation over a specific column in `mtcars`. Also, each function is named after the column over which it is supossed to do the transformation.

The `for` cycle selects the corresponding columns in `mtcars` using the functions names, and then replaces them with the modified versions, which are obtained applying the corresponding from the list.

## 21.4 For loops vs. functionals

### 21.4.1 Exercises

1. Read the documentation for `apply()`. In the 2d case, what two for loops does it generalise?

It generalises two nested for loops iterating over a two dimensions array.

2. Adapt `col_summary()` so that it only applies to numeric columns. You might want to start with an `is_numeric()` function that returns a logical vector that has a TRUE corresponding to each numeric column.

```{r}
is_col_numeric <- function(df) {
  out <- vector("logical", length(df))
  for (i in seq_along(df)) {
    out[[i]] <- class(df[[i]]) == "numeric"
  }
  out
}


col_summary <- function(df, fun) {
  subset_df <- df[is_col_numeric(df)]
  out <- vector("double", length(subset_df))
  for (i in seq_along(subset_df)) {
    out[i] <- fun(subset_df[[i]])
  }
  out
}

col_summary(iris, mean)
```

## 21.5 The map functions

### 21.5.3 Exercises

1. Write code that uses one of the map functions to:

  1. Compute the mean of every column in `mtcars`.
```{r}
map_dbl(mtcars, mean)
```
  
  2. Determine the type of each column in `nycflights13::flights`
```{r}
map_chr(mtcars, typeof)
```
  
  3. Compute the number of unique values in each column of `iris`
  
```{r}
map_dbl(iris, ~length(unique(.)))
```
  
  4. Generate 10 random normals for each of $\mu = -10$, $0$, $10$, and $100$
  
```{r}
means <- c(-10, 0, 10, 100)
map(means, ~rnorm(10, mean = .))
```
  
2. How can you create a single vector that for each column in a data frame indicates whether or not itÅfs a factor?

```{r}
map_lgl(mtcars, is_factor)
```

3. What happens when you use the map functions on vectors that aren't lists? What does `map(1:5, runif)` do? Why?

```{r}
map(1:5, runif)
```

It iterates over each element of the vector, regardless of whether it is an atomic vector or a list. In this case, it passes each element of `1:5` to the function `runif` as the first argument (`n`).

4. What does `map(-2:2, rnorm, n = 5)` do? Why? What does `map_dbl(-2:2, rnorm, n = 5)` do? Why?

```{r}
map(-2:2, rnorm, n = 5)
```

It passes each element of `-2:2` to `rnorm()` as the second argument (`mean`), since the first argument (`n`) is specified through `...`. Therefore, it creates a list of five vectors of length five, generated from normal distributions with means going from `-2` to `2`.

```{r}
map_dbl(-2:2, rnorm, n = 5)
```

This code tries to do the same, but it fails since `map_dbl()` tries to coerce the output to an atomic numeric vector, but that can not be done directly because each element of the result has length more than 1.

5. Rewrite `map(x, function(df) lm(mpg ~ wt, data = df))` to eliminate the anonymous function.

```
map(x, ~lm(mpg ~ wt, data = .))
```

## 21.9 Other patterns of for loops

### 21.9.3 Exercises

1. Implement your own version of `every()` using a for loop. Compare it with `purrr::every()`. What does purrr's version do that your version doesn't?

```{r}
x <- list(1:5, letters, list(10))

my_every <- function(x, fun) {
  output <- vector("logical", length(x))
  for (i in seq_along(x)) {
    output[[i]] <- fun(x[[i]])
  }
  all(output)
}

x %>% my_every(is_vector)
```

purrr version allows to pass additional arguments to the predicate function using `...`. 

2. Create an enhanced `col_summary()` that applies a summary function to every numeric column in a data frame

```{r}
is_col_numeric <- function(df) {
  map_lgl(df, ~class(.) == "numeric")
}


col_summary <- function(df, fun) {
  subset_df <- df[is_col_numeric(df)]
  
  map_dbl(subset_df, fun)
}

col_summary(iris, mean)
```

3. A possible base R equivalent of `col_summary()` is: 

```{r}
col_sum3 <- function(df, f) {
  is_num <- sapply(df, is.numeric)
  df_num <- df[, is_num]

  sapply(df_num, f)
}
```

But it has a number of bugs as illustrated with the following inputs:
```{r}
df <- tibble(
  x = 1:3, 
  y = 3:1,
  z = c("a", "b", "c")
)
# OK
col_sum3(df, mean)
# Has problems: don't always return numeric vector
col_sum3(df[1:2], mean)
col_sum3(df[1], mean)
col_sum3(df[0], mean)
```

In the case where `df` is subsetted by the empty vector `numeric(0)`, `sapply(df, is.numeric)` returns a list instead of a logical, causing the subsequent subsetting operation (`df[, is_num`) to fail (since we can not use a list to do subsetting, only a logical vector).

Also, even if the subsetting operation succeded in this case, the output wouldn't be a numeric vector, because `sapply(df_num, f)` returns a list when the result has length zero.

We can avoid both errors by replacing `sapply` with `map_lgl` and `map_dbl` in each call, respectively.
