---
title: "Chapter 24 - Exercises - R for Data Science"
author: "Francisco Yira Albornoz"
date: "13-12-2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
options(na.action = na.warn)

library(nycflights13)
library(lubridate)
```


## 24.2 Why are low quality diamonds more expensive?

### 24.2.3 Exercises

1. In the plot of `lcarat` vs. `lprice`, there are some bright vertical strips. What do they represent?

```{r}
diamonds2 <- diamonds %>% 
  filter(carat <= 2.5) %>% 
  mutate(lprice = log2(price), lcarat = log2(carat))

ggplot(diamonds2, aes(lcarat, lprice)) + 
  geom_hex(bins = 50)
```

```{r}
ggplot(diamonds2, aes(x = lcarat)) +
  geom_histogram(binwidth = 0.05)
```

It looks like bright areas in the first plot represent peaks or concentracions in the distribution of `lcarat` variable. Since data is asymetrically distributed around these concentrations, they may be caused by caused by "upward rounding" in data entry.

2. If `log(price) = a_0 + a_1 * log(carat)`, what does that say about the relationship between `price` and `carat`?

It means that `price` grows as `carat` grows following the relationship `e^(a_0) * carat^(a_1)`.

3. Extract the diamonds that have very high and very low residuals. Is there anything unusual about these diamonds? Are they particularly bad or good, or do you think these are pricing errors?

```{r}
mod_diamond2 <- lm(lprice ~ lcarat + color + cut + clarity, data = diamonds2)

diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamond2, "lresid2")

diamonds2 %>% 
  filter(abs(lresid2) > 1) %>% 
  add_predictions(mod_diamond2) %>% 
  mutate(pred = round(2 ^ pred)) %>% 
  select(price, pred, carat:table, x:z) %>% 
  arrange(price)
```

They mostly look like pricing errors. The prices predicted by the model are reasonable given all other variables. However, these low/high residuals could be a syntom of non-linear or interactive effects of some variables on diamonds price.

4. Does the final model, `mod_diamond2`, do a good job of predicting diamond prices? Would you trust it to tell you how much to spend if you were buying a diamond?

It appropriately captures the linear effects of the variables that affect the price, and since there are relatively few observations with large residuals, we could say that non-linear effects and interactions are not relevant in most cases. I would say the the final model sets a good benchmark of how much should we expect to pay for a diamond of certain characteristics.


## 24.3 What affects the number of daily flights?

### 24.3.3 Exercises

1. Use your Google sleuthing skills to brainstorm why there were fewer than expected flights on Jan 20, May 26, and Sep 1. (Hint: they all have the same explanation.) How would these days generalise to another year?

There were fewer flights in those days because they were close to Martin Luther King day, trinity Sunday and Labor day. The specific date in which these holidays fall depends on the weekdays of each year, so to generalise them to other years we need to keep that in mind.

2. What do the three days with high positive residuals represent? How would these days generalise to another year?

```{r}
daily <- flights %>% 
  mutate(date = make_date(year, month, day)) %>% 
  group_by(date) %>% 
  summarise(n = n()) %>% 
  mutate(wday = wday(date, label = TRUE))


term <- function(date) {
  cut(date,
      breaks = ymd(20130101, 20130605, 20130825, 20140101),
      labels = c("spring", "summer", "fall"))
}

daily <- daily %>% 
  mutate(term = term(date))

library(MASS)
mod <- MASS::rlm(n ~ wday * term, data = daily)

daily <- daily %>% 
  add_residuals(mod, "resid")

daily %>% 
  top_n(3, resid) 
```

These are the days after Thanksgiving (2013-11-28) and Christmas (2013-12-25). To generalise Thanksgiving to other years we should look for the fourth thursday of november. To generalise the peak after christmas we should simply look for the day after 25 Dec.

3. Create a new variable that splits the wday variable into terms, but only for Saturdays, i.e. it should have `Thurs`, `Fri`, but `Sat-summer`, `Sat-spring`, `Sat-fall`. How does this model compare with the model with every combination of wday and term?


```{r}
daily <- daily %>% 
  mutate(wday2 = ifelse(wday == "sab\\.",
                         yes = str_c(wday, term),
                         no = wday))

mod2 <- MASS::rlm(n ~ wday2, data = daily)

daily %>% 
  gather_residuals(all_terms = mod, sat_terms = mod2) %>% 
  ggplot(aes(x = date, y = resid, color = model)) +
  geom_line(alpha = 0.5) +
  geom_ref_line(h = 0)
```

It looks like the model with terms interactions in all weekdays performs better in summer, but worse in late-spring. 

4. Create a new `wday` variable that combines the day of week, term (for Saturdays), and public holidays. What do the residuals of that model look like?

```{r}
holidays <- ymd(20130101, 20130121, 20130218, 20130416, 20130512, 20130527, 20130616, 20130704, 20130902, 20131014, 20131111, 20131128, 20131129, 20131225)

daily <- daily %>% 
  mutate(wday3 = case_when(
    date %in% holidays ~ str_c(wday2, "_holiday"),
    TRUE ~ wday2
  ))

mod3 <- MASS::rlm(n ~ wday3, data = daily)

daily %>% 
  gather_residuals(without_holidays = mod2, with_holidays = mod3) %>% 
  ggplot(aes(x = date, y = resid, color = model)) +
  geom_line(alpha = 0.5) +
  geom_ref_line(h = 0)

```

The model estimates, in average, a positive and large effect due to holiday days. This helps to reduce residuals in some holidays which actually have more air traffic. However, in other days we see larger residuals since they correspond to holidays when air traffic doesn't increase so much.

5. What happens if you fit a day of week effect that varies by month (i.e. `n ~ wday * month`)? Why is this not very helpful?

```{r}
daily <- daily %>% 
  mutate(month = month(date))

mod4 <- MASS::rlm(n ~ wday * month(date), data = daily)

daily %>% 
  gather_residuals(mod_terms = mod, mod_months = mod4) %>% 
  ggplot(aes(x = date, y = resid, color = model)) +
  geom_line(alpha = 0.5) +
  geom_ref_line(h = 0)
```

In some periods the second model has larger residuals, this may be due to higher sensitivity to outliers. 

6. What would you expect the model `n ~ wday + ns(date, 5)` to look like? Knowing what you know about the data, why would you expect it to be not particularly effective?

```{r}
library(splines)
mod_sp <- MASS::rlm(n ~ wday * ns(date, 5), data = daily)

daily %>% 
  gather_residuals(mod_terms = mod, mod_splines = mod_sp) %>% 
  ggplot(aes(x = date, y = resid, color = model)) +
  geom_line(alpha = 0.5) +
  geom_ref_line(h = 0)
```

This new model is not clearly better than our original model with interactions between `wday` and `term`. A reason for this could be that it tries to fit smooth curves the data, but we see that there are abrupt changes in air traffic in some periods.

7. We hypothesised that people leaving on Sundays are more likely to be business travellers who need to be somewhere on Monday. Explore that hypothesis by seeing how it breaks down based on distance and time: if itÅfs true, youÅfd expect to see more Sunday evening flights to places that are far away.

```{r}
flights %>% 
  mutate(date = make_date(year, month, day),
         wday = wday(date, label = TRUE),
         time = make_datetime(year, month, day, dep_time %/% 100, dep_time %% 100),
         part_day = case_when(
           dep_time < 1200 ~ "AM",
           TRUE ~ "PM"
         ),
         wday_and_part = str_c(wday, part_day, sep = "_")) %>% 
  ggplot(aes(x = wday_and_part, y = distance)) +
  geom_boxplot()
```

It looks like the median distance is actually higher in Sunday morning (not Sunday evening).

An alternative way to test this hypothesis could be a plot of distance traveled (in Sunday) against hours left to midnight.

```{r}
flights %>% 
  mutate(date = make_date(year, month, day),
         wday = wday(date, label = TRUE),
         hour = dep_time %/% 100) %>% 
  filter(wday == "dom\\.") %>% 
  ggplot(aes(x = as.factor(hour), y = distance)) +
  geom_boxplot()
```

We see a subtle increase in distance traveled around 17 and 19 hr. of Sunday.

8. ItÅfs a little frustrating that Sunday and Saturday are on separate ends of the plot. Write a small function to set the levels of the factor so that the week starts on Monday.

```{r}
order_wdays <- function(data) {
  data %>%
    mutate(wday = factor(
      wday,
      levels = c(
        "lun\\.",
        "mar\\.",
        "mie\\.",
        "jue\\.",
        "vie\\.",
        "sab\\.",
        "dom\\."
      )
    ))
}

daily %>% order_wdays() %>% 
  ggplot(aes(x = wday)) +
  geom_bar()
```
