---
title: "Chapter 27 - Exercises - R for Data Science"
author: "Francisco Yira Albornoz"
date: "09-02-2019"
output: html_document
---

## 27.3 Text formatting with R Markdown

### 27.3.1 Exercises

2. Using the R Markdown quick reference, figure out how to:

    1. Add a footnote.
    2. Add a horizontal rule.
    3. Add a block quote.
  
> This is a block quote

-----
This is a horizontal rule

This is a footnote [^1] 

[^1]: Here is the footnote. 

3. Copy and paste the contents of `diamond-sizes.Rmd` from https://github.com/hadley/r4ds/tree/master/rmarkdown in to a local R markdown document. Check that you can run it, then add text after the frequency polygon that describes its most striking features.

## 27.4 Code chunks

### 27.4.7 Exercises

(answers in `diamond-sizes.Rmd` inside this repo)

1. Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting `echo = FALSE` on each chunk, set a global option.

2. Add a section that describes the largest 20 diamonds, including a table that displays their most important attributes.

3. Modify `diamonds-sizes.Rmd` to use `comma()` to produce nicely formatted output. Also include the percentage of diamonds that are larger than 2.5 carats.

4. Set up a network of chunks where `d` depends on `c` and `b`, and both `b` and `c` depend on `a`. Have each chunk print `lubridate::now()`, set `cache = TRUE`, then verify your understanding of caching.

```{r a}
lubridate::now()
```

```{r b, dependson="a"}
lubridate::now()
```

```{r c, dependson="a"}
lubridate::now()
```

```{r d, dependson=c("b", "c"), cache=TRUE}
lubridate::now()
```